Multidimensional signals and analytic flexibility: Estimating degrees of freedom in human speech analyses

### 2021-08-20 RR_manuscript/Draft_RR.Rmd
minor edits based on reviews

To further illuminate the question of whether speakers modify speech to signal atypical referents, and overcome some of the limitations of previous work, thirty native German speakers were recorded in a production study while interacting with a confederate (one of the experimenters) in a referential game.
The participants had to verbally instruct the confederate to select a specified target object out of four objects presented on a screen.
The subject and confederate were seated at the opposite sides of a table, each facing one of two computer screens.
The participant and the experimenter could not see each other nor each others' screens.
Figure \@ref(fig:procedural) shows the experimental procedure time-line.
After a familiarization phase, the subject first saw four colored objects in the top left, top right, bottom left, and bottom right corners of the screen.
One of the objects served as the target, another as the competitor, and the remaining two objects served as distractors.
Objects were referred to using noun phrases consisting of an adjective modifier denoting color and a modified object (e.g. *Gelbe Zitrone* 'yellow lemon', *Rote Gurke* 'red cucumber', *Rote Socken* 'red socks').

@@

The two sentence prompts were used to create a focus contrast between the competitor and the target object. Focused units denote the set of all (contextually relevant) alternatives [e.g. @rooth1992theory].Concretely, a focus contrast encodes information that is assumed to be contrary to the presuppositions of the interlocutor. 
For instance, if the competitor and target objects differ but their color did not (e.g. *yellow banana* vs *yellow tomato*), the noun is in focus (the Noun Focus condition, NF).
If the objects are the same but differ in color (e.g. *yellow banana* vs *blue banana*), the color adjective is in focus (the Adjective Focus condition, AF).
If both the color and the object differ (e.g. *yellow banana* vs *blue tomato*), the the whole noun phrase is in focus (the Adjective/Noun Focus condition, ANF).
The NF condition constituted the experimentally relevant condition, while the AF and ANF conditions acted as fillers.
Crucially, the color-object combinations in the Noun Focus (NF) condition were manipulated with respect to their typicality.
The combinations were either typical (e.g. *orange mandarin*), medium typical (e.g. *green tomato*), or atypical (e.g. *yellow cherry*), as established by a norming study that was conducted prior to the production experiment just described.^[A detailed description of the norming and production studies from the *Prosodic encoding of redundant referring expressions* project, which was given to the analysts with the data set, can be found in `methods_norm_prod.pdf` at (temporary link) <https://osf.io/5agn9/?view_only=3429acfd8eae4f0aafe2e17ad5b0983f>]
Each subject produced 15 critical trials (NF condition).
Each trial was repeated twice, yielding a total of 30 trials per participant and a grand-total of 900 ($15 \times 2 \times 30$ participants) spoken utterances.

For the present study, N analysis teams will receive access to the entire data set generated by the production study and will be instructed to answer the following research question: *Do speakers acoustically modify utterances to signal atypical word combinations?*

@@

Since we will obtain scores for each member in the analytic teams, we will calculate the team mean and standard deviation for each predictor.
These will be entered in the model formula as a measurement-error term (`me(mean, sd)` in brms).

>>>>

To further illuminate the question of whether speakers modify speech to signal atypical referents, and overcome some of the limitations of previous work, thirty native German speakers were recorded in a production study while interacting with a confederate (one of the experimenters) in a referential game, following expermintal procedures typical of the field.
The participants had to verbally instruct the confederate to select a specified target object out of four objects presented on a screen.
The subject and confederate were seated at the opposite sides of a table, each facing one of two computer screens.
The participant and the experimenter could not see each other nor each others' screens.
Figure \@ref(fig:procedural) shows the experimental procedure time-line.
After a familiarization phase, the subject first saw four colored objects in the top left, top right, bottom left, and bottom right corners of the screen.
One of the objects served as the target, another as the competitor, and the remaining two objects served as distractors.
Objects were referred to using noun phrases consisting of an adjective modifier denoting color and a modified object (e.g. *Gelbe Zitrone* 'yellow lemon', *Rote Gurke* 'red cucumber', *Rote Socken* 'red socks').

@@

The two sentence prompts were used to create a focus contrast between the competitor and the target object.
Focused units denote the set of all (contextually relevant) alternatives [e.g. @rooth1992theory].
Concretely, a focus contrast marks one of more elements in a sentence as prominent, by different linguistic means depending on the language.
For instance, if the competitor and target objects differ but their color did not (e.g. *yellow banana* vs *yellow tomato*), the noun is said to be in focus (the Noun Focus condition, NF).
If the objects are the same but differ in color (e.g. *yellow banana* vs *blue banana*), the color adjective is in focus (the Adjective Focus condition, AF).
If both the color and the object differ (e.g. *yellow banana* vs *blue tomato*), the the whole noun phrase is in focus (the Adjective/Noun Focus condition, ANF).
The NF condition constituted the experimentally relevant condition, while the AF and ANF conditions acted as fillers.
Crucially, the color-object combinations in the Noun Focus (NF) condition were manipulated with respect to their typicality.
The combinations were either typical (e.g. *orange mandarin*), medium typical (e.g. *green tomato*), or atypical (e.g. *yellow cherry*), as established by a norming study that was conducted prior to the production experiment just described.^[A detailed description of the norming and production studies from the *Prosodic encoding of redundant referring expressions* project, which was given to the analysts with the data set, can be found in `methods_norm_prod.pdf` at (temporary link) <https://osf.io/5agn9/?view_only=3429acfd8eae4f0aafe2e17ad5b0983f>]
Each subject produced 15 critical trials (NF condition).
Each trial was repeated twice, yielding a total of 30 trials per participant and a grand-total of 900 ($15 \times 2 \times 30$ participants) spoken utterances.

For the present study, N analysis teams will receive access to the entire data set generated by the production study.
The data set is constituded by audio recordings and annotation files in a format that is typical for the field.
The teams will be instructed to answer the following research question, using the provided data set: *Do speakers acoustically modify utterances to signal atypical word combinations?*

@@

To obtain an aggregated research experience score and initial belief score for each team based on the members' individual scores, we will calculate the mean and standard deviation of these predictors for each team.
These will be entered in the model formula as a measurement-error term (`me(mean, sd)` in brms).
The expedient of using a measurement-error term (which includes the teams' standard deviation) ensures information about within-team variance is not lost (which would be the case if including the mean only).

### 2021-08-20 RR_manuscript/Draft_RR.tex
minor edits based on reviews

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

@@

\hypertarget{researcher-degrees-of-freedom}{%
\subsection{Researcher degrees of freedom}\label{researcher-degrees-of-freedom}}

@@

\hypertarget{crowdsourcing-alternative-analyses}{%
\subsection{Crowdsourcing alternative analyses}\label{crowdsourcing-alternative-analyses}}

@@

\hypertarget{s:operspeech}{%
\subsection{Operationalizing speech}\label{s:operspeech}}

@@

\begin{enumerate}
\def\labelenumi{(\arabic{enumi})}
\tightlist
\item
  ``I can't bear another meeting on Zoom.''
  \vspace{1em}
\end{enumerate}

@@

From a phonetic viewpoint, linguistically relevant units are inherently multidimensional and dynamic: they consist of clusters of parameters that are modulated over time.
The acoustic parameters of units are usually asynchronous, i.e.~they appear at different time points in the unfolding signal, and overlap with parameters of other units \citep[e.g.][]{jongman2000acoustic, lisker1986voicing, summerfield1981articulatory, winter2014spoken}.
A classical example is the distinction between voiced and voiceless stops in English (i.e.~/b/ and /p/ in \emph{bear} vs \emph{pear}).
This contrast is manifested by many acoustic features which can differ depending on several factors, such as position of the consonant in the word and surrounding sounds \citep{lisker1977rapid}.
Furthermore, correlates of the contrast can even be found away from the consonant, in temporally distant speech units.
For example, the initial /l/ of the English words \emph{led} and \emph{let} is affected by the voicing of the final consonant (/d, t/) \citep{hawkins2004influence}.

@@

\begin{figure}
\includegraphics[width=1\linewidth]{/Users/ste/repos/many-speech/many_analyses/figs/ForkingPaths} \caption{Illustration of the analytic flexibility associated with acoustic analyses. (A) An example of multiple possible and justifiable decisions when comparing two utterances; (B) Waveform and fundamental frequency (f0) track of the utterances \emph{I can't bear ANOTHER meeting on Zoom} and \emph{I can't bear another meeting on ZOOM}. The green boxes mark the word \emph{another} in both sentences; (C) Spectrogram and f0 track of the word \emph{another}, exemplifying possible operationalizations of differences in f0.}\label{fig:forkingPaths}
\end{figure}

@@

\hypertarget{s:dataset}{%
\subsection{The data set: The acoustic properties of atypical modifiers}\label{s:dataset}}

@@

To further illuminate the question of whether speakers modify speech to signal atypical referents, and overcome some of the limitations of previous work, thirty native German speakers were recorded in a production study while interacting with a confederate (one of the experimenters) in a referential game.
The participants had to verbally instruct the confederate to select a specified target object out of four objects presented on a screen.
The subject and confederate were seated at the opposite sides of a table, each facing one of two computer screens.
The participant and the experimenter could not see each other nor each others' screens.
Figure \ref{fig:procedural} shows the experimental procedure time-line.
After a familiarization phase, the subject first saw four colored objects in the top left, top right, bottom left, and bottom right corners of the screen.
One of the objects served as the target, another as the competitor, and the remaining two objects served as distractors.
Objects were referred to using noun phrases consisting of an adjective modifier denoting color and a modified object (e.g.~\emph{Gelbe Zitrone} `yellow lemon', \emph{Rote Gurke} `red cucumber', \emph{Rote Socken} `red socks').

\begin{figure}
\includegraphics[width=1\linewidth]{/Users/ste/repos/many-speech/many_analyses/figs/procedure_plot} \caption{Experimental procedure. The upper row illustrates the trial sequence for the speaker (participant) and the lower row illustrates the trial sequence for the confederate. After a preview of 1500ms the speaker sees an arrow indicating one of the referent (b). Reading the orthographic instructions out loud, the speaker gives the confederate verbal instructions onto which referent they should drag the cube (c). The confederate, in turn, drags the black cube onto the target referent (d). Both the arrow and the orthographic instruction disappear from the speaker's screen and a new referent is indicated by an arrow on the same display alongside a new orthographic instruction (e). The speaker gives the confederate verbal instructions (f) which the confederate follows by dragging the cube onto the next referent (g).}\label{fig:procedural}
\end{figure}

@@

The two sentence prompts were used to create a focus contrast between the competitor and the target object. Focused units denote the set of all (contextually relevant) alternatives \citep[e.g.][]{rooth1992theory}.Concretely, a focus contrast encodes information that is assumed to be contrary to the presuppositions of the interlocutor.
For instance, if the competitor and target objects differ but their color did not (e.g.~\emph{yellow banana} vs \emph{yellow tomato}), the noun is in focus (the Noun Focus condition, NF).
If the objects are the same but differ in color (e.g.~\emph{yellow banana} vs \emph{blue banana}), the color adjective is in focus (the Adjective Focus condition, AF).
If both the color and the object differ (e.g.~\emph{yellow banana} vs \emph{blue tomato}), the the whole noun phrase is in focus (the Adjective/Noun Focus condition, ANF).
The NF condition constituted the experimentally relevant condition, while the AF and ANF conditions acted as fillers.
Crucially, the color-object combinations in the Noun Focus (NF) condition were manipulated with respect to their typicality.
The combinations were either typical (e.g.~\emph{orange mandarin}), medium typical (e.g.~\emph{green tomato}), or atypical (e.g.~\emph{yellow cherry}), as established by a norming study that was conducted prior to the production experiment just described.\footnote{A detailed description of the norming and production studies from the \emph{Prosodic encoding of redundant referring expressions} project, which was given to the analysts with the data set, can be found in \texttt{methods\_norm\_prod.pdf} at (temporary link) \url{https://osf.io/5agn9/?view_only=3429acfd8eae4f0aafe2e17ad5b0983f}}
Each subject produced 15 critical trials (NF condition).
Each trial was repeated twice, yielding a total of 30 trials per participant and a grand-total of 900 (\(15 \times 2 \times 30\) participants) spoken utterances.

For the present study, N analysis teams will receive access to the entire data set generated by the production study and will be instructed to answer the following research question: \emph{Do speakers acoustically modify utterances to signal atypical word combinations?}

\hypertarget{methods}{%
\section{Methods}\label{methods}}

@@

\hypertarget{phase-1-recruitment-of-analysts-and-initial-survey}{%
\subsection{Phase 1: Recruitment of analysts and initial survey}\label{phase-1-recruitment-of-analysts-and-initial-survey}}

@@

\hypertarget{phase-2-primary-data-analyses}{%
\subsection{Phase 2: Primary Data Analyses}\label{phase-2-primary-data-analyses}}

@@

\hypertarget{phase-3-peer-review-of-analyses}{%
\subsection{Phase 3: Peer Review of Analyses}\label{phase-3-peer-review-of-analyses}}

@@

\hypertarget{phase-4-evaluating-variation}{%
\subsection{Phase 4: Evaluating variation}\label{phase-4-evaluating-variation}}

@@

\hypertarget{descriptive-statistics}{%
\subsubsection{Descriptive statistics}\label{descriptive-statistics}}

@@

\hypertarget{s:meta-est}{%
\subsubsection{Meta-analytic estimation}\label{s:meta-est}}

@@

\hypertarget{anares-preds}{%
\subsubsection{Analytic and researcher-related predictors affecting effect sizes}\label{anares-preds}}

@@

Since we will obtain scores for each member in the analytic teams, we will calculate the team mean and standard deviation for each predictor.
These will be entered in the model formula as a measurement-error term (\texttt{me(mean,\ sd)} in brms).

@@

\hypertarget{ana-archive}{%
\subsubsection{Data management}\label{ana-archive}}

@@

\hypertarget{phase-5-collaborative-write-up-of-manuscript}{%
\subsection{Phase 5: Collaborative Write-Up of Manuscript}\label{phase-5-collaborative-write-up-of-manuscript}}

@@

\hypertarget{author-contributions}{%
\section{Author contributions}\label{author-contributions}}

@@

\hypertarget{conflicts-of-interest}{%
\section{Conflicts of interest}\label{conflicts-of-interest}}

@@

\hypertarget{glossary}{%
\section{Glossary}\label{glossary}}

>>>>

\section{Introduction}\label{introduction}

@@

\subsection{Researcher degrees of
freedom}\label{researcher-degrees-of-freedom}

@@

\subsection{Crowdsourcing alternative
analyses}\label{crowdsourcing-alternative-analyses}

@@

\hypertarget{s:operspeech}{\subsection{Operationalizing
speech}\label{s:operspeech}}

@@

\vspace{1em} (1) ``I can't bear another meeting on Zoom.'' \vspace{1em}

@@

From a phonetic viewpoint, linguistically relevant units are inherently
multidimensional and dynamic: they consist of clusters of parameters
that are modulated over time. The acoustic parameters of units are
usually asynchronous, i.e.~they appear at different time points in the
unfolding signal, and overlap with parameters of other units
\citep[e.g.][]{jongman2000acoustic, lisker1986voicing, summerfield1981articulatory, winter2014spoken}.
A classical example is the distinction between voiced and voiceless
stops in English (i.e. /b/ and /p/ in \emph{bear} vs \emph{pear}). This
contrast is manifested by many acoustic features which can differ
depending on several factors, such as position of the consonant in the
word and surrounding sounds \citep{lisker1977rapid}. Furthermore,
correlates of the contrast can even be found away from the consonant, in
temporally distant speech units. For example, the initial /l/ of the
English words \emph{led} and \emph{let} is affected by the voicing of
the final consonant (/d, t/) \citep{hawkins2004influence}.

@@

\begin{figure}
\includegraphics[width=1\linewidth]{/homes/s.coretta/repos/many_analyses/figs/ForkingPaths} \caption{Illustration of the analytic flexibility associated
with acoustic analyses. (A) An example of multiple possible and
justifiable decisions when comparing two utterances; (B) Waveform and
fundamental frequency (f0) track of the utterances \emph{I can't bear
ANOTHER meeting on Zoom} and \emph{I can't bear another meeting on
ZOOM}. The green boxes mark the word \emph{another} in both sentences;
(C) Spectrogram and f0 track of the word \emph{another}, exemplifying
possible operationalizations of differences in f0.}\label{fig:forkingPaths}
\end{figure}

@@

\hypertarget{s:dataset}{\subsection{The data set: The acoustic
properties of atypical modifiers}\label{s:dataset}}

@@

To further illuminate the question of whether speakers modify speech to
signal atypical referents, and overcome some of the limitations of
previous work, thirty native German speakers were recorded in a
production study while interacting with a confederate (one of the
experimenters) in a referential game, following expermintal procedures
typical of the field. The participants had to verbally instruct the
confederate to select a specified target object out of four objects
presented on a screen. The subject and confederate were seated at the
opposite sides of a table, each facing one of two computer screens. The
participant and the experimenter could not see each other nor each
others' screens. Figure \ref{fig:procedural} shows the experimental
procedure time-line. After a familiarization phase, the subject first
saw four colored objects in the top left, top right, bottom left, and
bottom right corners of the screen. One of the objects served as the
target, another as the competitor, and the remaining two objects served
as distractors. Objects were referred to using noun phrases consisting
of an adjective modifier denoting color and a modified object (e.g.
\emph{Gelbe Zitrone} `yellow lemon', \emph{Rote Gurke} `red cucumber',
\emph{Rote Socken} `red socks').

\begin{figure}
\includegraphics[width=1\linewidth]{/homes/s.coretta/repos/many_analyses/figs/procedure_plot} \caption{Experimental procedure. The upper row illustrates the
trial sequence for the speaker (participant) and the lower row
illustrates the trial sequence for the confederate. After a preview of
1500ms the speaker sees an arrow indicating one of the referent (b).
Reading the orthographic instructions out loud, the speaker gives the
confederate verbal instructions onto which referent they should drag the
cube (c). The confederate, in turn, drags the black cube onto the target
referent (d). Both the arrow and the orthographic instruction disappear
from the speaker's screen and a new referent is indicated by an arrow on
the same display alongside a new orthographic instruction (e). The
speaker gives the confederate verbal instructions (f) which the
confederate follows by dragging the cube onto the next referent (g).}\label{fig:procedural}
\end{figure}

@@

The two sentence prompts were used to create a focus contrast between
the competitor and the target object. Focused units denote the set of
all (contextually relevant) alternatives
\citep[e.g.][]{rooth1992theory}. Concretely, a focus contrast marks one
of more elements in a sentence as prominent, by different linguistic
means depending on the language. For instance, if the competitor and
target objects differ but their color did not (e.g. \emph{yellow banana}
vs \emph{yellow tomato}), the noun is said to be in focus (the Noun
Focus condition, NF). If the objects are the same but differ in color
(e.g. \emph{yellow banana} vs \emph{blue banana}), the color adjective
is in focus (the Adjective Focus condition, AF). If both the color and
the object differ (e.g. \emph{yellow banana} vs \emph{blue tomato}), the
the whole noun phrase is in focus (the Adjective/Noun Focus condition,
ANF). The NF condition constituted the experimentally relevant
condition, while the AF and ANF conditions acted as fillers. Crucially,
the color-object combinations in the Noun Focus (NF) condition were
manipulated with respect to their typicality. The combinations were
either typical (e.g. \emph{orange mandarin}), medium typical (e.g.
\emph{green tomato}), or atypical (e.g. \emph{yellow cherry}), as
established by a norming study that was conducted prior to the
production experiment just described.\footnote{A detailed description of
  the norming and production studies from the \emph{Prosodic encoding of
  redundant referring expressions} project, which was given to the
  analysts with the data set, can be found in
  \texttt{methods\_norm\_prod.pdf} at (temporary link)
  \url{https://osf.io/5agn9/?view_only=3429acfd8eae4f0aafe2e17ad5b0983f}}
Each subject produced 15 critical trials (NF condition). Each trial was
repeated twice, yielding a total of 30 trials per participant and a
grand-total of 900 (\(15 \times 2 \times 30\) participants) spoken
utterances.

For the present study, N analysis teams will receive access to the
entire data set generated by the production study. The data set is
constituded by audio recordings and annotation files in a format that is
typical for the field. The teams will be instructed to answer the
following research question, using the provided data set: \emph{Do
speakers acoustically modify utterances to signal atypical word
combinations?}

\section{Methods}\label{methods}

@@

\subsection{Phase 1: Recruitment of analysts and initial
survey}\label{phase-1-recruitment-of-analysts-and-initial-survey}

@@

\subsection{Phase 2: Primary Data
Analyses}\label{phase-2-primary-data-analyses}

@@

\subsection{Phase 3: Peer Review of
Analyses}\label{phase-3-peer-review-of-analyses}

@@

\subsection{Phase 4: Evaluating
variation}\label{phase-4-evaluating-variation}

@@

\subsubsection{Descriptive statistics}\label{descriptive-statistics}

@@

\subsubsection{Meta-analytic estimation}\label{s:meta-est}

@@

\subsubsection{Analytic and researcher-related predictors affecting
effect sizes}\label{anares-preds}

@@

To obtain an aggregated research experience score and initial belief
score for each team based on the members' individual scores, we will
calculate the mean and standard deviation of these predictors for each
team. These will be entered in the model formula as a measurement-error
term (\texttt{me(mean,\ sd)} in brms). The expedient of using a
measurement-error term (which includes the teams' standard deviation)
ensures information about within-team variance is not lost (which would
be the case if including the mean only).

@@

\subsubsection{Data management}\label{ana-archive}

@@

\subsection{Phase 5: Collaborative Write-Up of
Manuscript}\label{phase-5-collaborative-write-up-of-manuscript}

@@

\section{Author contributions}\label{author-contributions}

@@

\section{Conflicts of interest}\label{conflicts-of-interest}

@@

\section{Glossary}\label{glossary}
